The Development of a Collaborative Tool to Teach Debugging

### 2020-11-28 paper/paper.tex
finished draft of tool writeups

\subsection{Project Components}

@@

\begin{enumerate}
\item A frontend web app built using React (\ref{react}) that presents a debugging
  interface to the end user.
\item A distributed backend managed by Kubernetes (\ref{k8s}), split into three parts:
  \begin{enumerate}
  \item A pod for each debugging instance which runs the rr debugger
    (\ref{rr}).  These communicate directly with users through
    WebSocket server pods. (\ref{socketio}).
  \item Pods running a frontend server written in Node which works in
    tandem with an API server created using Flask (\ref{flask/node}).
    The API server manages creation and destruction of debug sessions,
    as well as authentication.
  \item A MongoDB database.
  \end{enumerate}
\end{enumerate}

@@

@@

To speed communication, the collaborative debugger uses WebSockets to
directly connect web clients and the pods running rr.  Socket.IO is a
library that extends WebSockets.  It provides backup incase a
WebSocket connection cannot be established, enables automatic
reconnection and disconnection detection, and adds support for
namespaces \cite{socketio}.  The collaborative debugger uses the
standard Javascript implementation of Socket.IO on the client side.
Messages are passed through a server to individual debugging pods,
both of which use the Python implementation of Socket.IO,
python-socket.io \cite{python_socketio}.

\subsubsection{Flask and Node}\label{flask/node}

\subsubsection{React and Xterm.js}\label{react}

>>>>

\subsection{Tools Used}

@@

\begin{enumerate}
\item A frontend web app built using React (\ref{react}) that presents a debugging
  interface to the end user.
\item A distributed backend managed by Kubernetes (\ref{k8s}), split into three parts:
  \begin{enumerate}
  \item A pod for each debugging instance which runs the rr debugger
    (\ref{rr}).  These communicate directly with users through
    WebSocket server pods. (\ref{socketio}).
  \item Pods running a frontend server written in Node.js which works
    in tandem with an API server created using Flask
    (\ref{flask/node}).  The API server manages creation and
    destruction of debug sessions, as well as authentication.
  \item A MongoDB (\ref{mongodb}) database.
  \end{enumerate}
\end{enumerate}

@@

In order to ``support the development of systems which use the
debugger as just one small component of a larger system'', GDB
provides a machine-oriented interface called GDB/MI \cite{gdbman}. rr
supports interaction through GDB/MI, and using the interface was a
natural choice for the collaborative debugger.  In addition to being
far easier to interact with from within a program, the structured,
machine-friendly output of GDB/MI lends itself in particular to future
development of visualization aids in the collaborative debugger.
\par
To parse rr output into Python dictionaries and to easily control rr
as a subprocess, pygdbmi \cite{pygdbmi} is used in each debugging pod.
pygdbmi's abstraction simplifies programatically controlling rr.  A
pod can receive a command from the client, pass it to rr, and respond
without having to deal with parsing GDB/MI output or managing the rr
process.

@@

To speed communication, the collaborative debugger uses WebSockets to
directly connect web clients and the pods running rr.  Socket.IO is a
library that extends WebSockets.  It provides backup incase a
WebSocket connection cannot be established, enables automatic
reconnection and disconnection detection, and adds support for
namespaces \cite{socketio}.  The collaborative debugger uses the
standard JavaScript implementation of Socket.IO on the client side.
Messages are passed through a server to individual debugging pods,
both of which use the Python implementation of Socket.IO,
python-socket.io \cite{python_socketio}.

\subsubsection{MongoDB}\label{mongodb}

The collaborative debugger uses a database to store information about
users, pods, and example debugging sessions.  Due to it's speed of
deployment and natural interaction with the object-oriented languages
used to create the project, MongoDB was chosen as database software
\cite{mongodb}.

\subsubsection{Flask and Node.js}\label{flask/node}

The primary server for the collaborative debugger is split into two
sections: a simple Node.js \cite{node} server that serves the frontend
webapp, and an API server created using Flask \cite{flask}.  While in
development, the builtin React (see next section) development server
is used to serve the frontend.  This makes debugging the frontend far
easier.
\par

An API server is necessary to authenticate users and to provide a
means to create/delete debugging sessions.  Since the rest of the
backend was created using Python, Flask was chosen to create the API
server.  Flask is a lightweight web application framework which lends
itself perfectly to interacting with the Python MongoDB and Kubernetes
APIs.

\subsubsection{React}\label{react}

React is JavaScript library that simplifies creating user interfaces
and managing state \cite{react}.  React's state management is of
particular importance to the collaborative debugger's frontend.  State
constantly changes as users create/delete debugging sessions, join
existing sessions, and communicate with rr.  React allows classes to
encapsulate components such as a list of existing debugging sessions,
a view of the current program's source code, and the terminal
interface with rr.  Instances of these classes maintain state
and update efficiently.
\par
The frontend makes extensive use of JSX, syntax which allows the
inclusion of segments of HTML code within a React app written in
JavaScript.  This makes it easy for each component of the one-page
webapp to hide/show subcomponents as state changes.

\subsubsection{Monaco and Xterm.js}\label{xtermjs/monaco}

After joining or creating a debugging session, users spend most of
their time interacting collaboratively with rr.  Their primary
interface to rr is through Xterm.js, a frontend component that makes
it easy to emulate terminal behavior in the browser \cite{xtermjs}.
With a few control methods, it is simple to provide a terminal
interface to rr that is virtually indistinguishable from a local
session.  By using the Xterm.js based interface, students can learn to
use rr (and by extension gdb) collaboratively, and directly translate
that knowledge to individual work.
\par

In addition to the terminal interface, the frontend shows a view of
the current source file being debugged.  The Monaco Editor
\cite{monaco} is used to display this source view.  Though more
complex than is strictly necessary to display code, Monaco makes it
easy to format and syntax-highlight.  Using Monaco also simplifies the
future addition of editing source code, should the need arise.
React's state management allows updating text in the editor as
efficiently as possible.
